Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans

### 2020-05-20 paper_PLOS/paper_PLOS_draft.Rmd
first draft

---
title: Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Paenungulates
author:
  - name: Juan Manuel Vazquez
    email: juanvazquez@uchicago.edu
    affiliation: University of Chicago
    corresponding: juanvazquez@uchicago.edu
  - name: Vincent J Lynch
    email: vjlynch@buffalo.edu
    affiliation: University at Buffalo, SUNY
address:
  - code: University of Chicago
    address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
  - code: University at Buffalo, SUNY
    address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
bibliography: rbhb_paenungulates.bib
output: 
  rticles::plos_article: 
    includes:
      in_header: preamble.tex
  workflowr::wflow_html: default
  word_document: default
csl: plos.csl
editor_options:
  chunk_output_type: console
---

@@

```{r Libraries and functions, include = F}
knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=5, 
                      fig.retina = T)

@@

anage <- read_tsv("../data/AnAge/anage_build14.txt", col_types = list("Weaning (days)"=col_number(), "Weaning weight (g)"=col_number(), "References" = col_character())) %>% 
  select(Genus, Species,"Maximum longevity (yrs)", "Adult weight (g)") %>% 
  unite("label", Genus, Species, sep=".") %>% 
  mutate(lnSize.anage = sapply(
    `Adult weight (g)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  ),
  Lifespan = sapply(
    `Maximum longevity (yrs)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  )
  ) %>% 
  select(label, lnSize.anage, Lifespan)
```
```{r Query List, include=F}
UP000005640.full <- UP000005640 %>% 
  filter(!is.na(Name))
UP000005640.genenames <- unique(UP000005640.full$Name)
query.binTable <- list(AvA = gene_list, UP000005640 = UP000005640.genenames) %>% 
  binTableFromList() %>% 
  left_join(., UP000005640, by=c("Gene"="Name")) %>% 
  mutate(
    "C.ORF" = str_detect(Gene, "^C[0-9XY]{1,2}orf\\d+"),  # No C-Orfs
    "LOC" = str_detect(Gene, "^LOC\\d+"),   # No uncharacterized loci
    "LINC" = str_detect(Gene, "^LINC\\d+"),   # No Long "Non-Coding" RNA
    "HLA" = str_detect(Gene, "^HLA|^HCP5|^CD74"), # No HLAs
    "Histone" = str_detect(Gene, "^H[1-4]|^HIST\\d"), # No Histones
    "ZincFinger" = str_detect(Gene, "^ZNF\\d+"),   # No Zinc Fingers
    "Riboprotein" = str_detect(Gene, "^RP[LS]\\d+"),   # No Riboproteins
    "OlfactoryReceptor" = str_detect(Gene, "^OR\\d+"),   # No Olfactory Receptors
    "Uncharacterized" = str_detect(header, "[uU]ncharacterized"),  # No Uncharacterized proteins
    "Putative" = str_detect(header, "[pP]utative"),  # No putative proteins
    "Viral" = str_detect(header, "[pP]olyprotein|ERVK|HERV"),  # No viral proteins
    "Fragment" = str_detect(header, "[Ff]ragment")  # No fragment proteins
  )
```
```{r Body Size Data: BSize only, include=F}
bsize.innerNodes <- tribble(
  ~"Node",                                                     ~"label",
  MRCA(bsize.tree, "Loxodonta.africana", "Loxodonta.cyclotis"), "Loxodontini",
  MRCA(bsize.tree, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"), "Loxodona",
  MRCA(bsize.tree, "Loxodonta.africana", "Elephas.maximus"), "Elephantidae",
  MRCA(bsize.tree, "Elephas.maximus", "Mammuthus.primigenius"), "Elephantina",
  MRCA(bsize.tree, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
  MRCA(bsize.tree, "Loxodonta.africana", "Mammut.americanum"), "Proboscidae",
  MRCA(bsize.tree, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
  MRCA(bsize.tree, "Loxodonta.africana", "Procavia.capensis"), "Paenungulata",
  MRCA(bsize.tree, "Loxodonta.africana", "Orycteropus.afer"), "Pseudoungulata",
  MRCA(bsize.tree, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
  MRCA(bsize.tree, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosorcida",
  MRCA(bsize.tree, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
  MRCA(bsize.tree, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), "Xenarthra",
  MRCA(bsize.tree, "Dasypus.novemcinctus", "Loxodonta.africana"), "Atlantogenata"
)
bsize.tree %<>% as_tibble()

@@

size.scale <- bsize.tree %>% as_tibble %>% pull(lnSize) %>% unique %>% as.numeric() %>% sort
size.scale.min <- size.scale[1]
size.scale.max <- size.scale[length(size.scale)]
get.size.color <- function(n, scale = size.scale){
  scale.colors = viridis::viridis(n=length(scale)) %>% 
    set_names(., as.character(scale))
  scale.colors[as.character(n)]
}
bsize.Atlantogenata.nodes <- list(
  "Loxodonta.africana"= bsize.Atlantogenata.table %>% filter(label=="Loxodonta africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= bsize.Atlantogenata.table %>% filter(label=="Trichechus manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= bsize.Atlantogenata.table %>% filter(label=="Procavia capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= bsize.Atlantogenata.table %>% filter(label=="Orycteropus afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= bsize.Atlantogenata.table %>% filter(label=="Choloepus hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = bsize.Atlantogenata.table %>% filter(label=="Dasypus novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = bsize.Atlantogenata.table %>% filter(label=="Chrysochloris asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = bsize.Atlantogenata.table %>% filter(label=="Echinops telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = bsize.Atlantogenata.table %>% filter(label=="Elephantulus edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = bsize.Atlantogenata.table %>% filter(label=="Loxodonta cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = bsize.Atlantogenata.table %>% filter(label=="Palaeoloxodon antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = bsize.Atlantogenata.table %>% filter(label=="Elephas maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = bsize.Atlantogenata.table %>% filter(label=="Mammut americanum") %>% pull(node) %>% unique,
  "Loxodontini" = bsize.Atlantogenata.table %>% filter(label=="Loxodontini") %>% pull(node) %>% unique,
  "Loxodona" = bsize.Atlantogenata.table %>% filter(label=="Loxodona") %>% pull(node) %>% unique,
  "Elephantidae" = bsize.Atlantogenata.table %>% filter(label=="Elephantidae") %>% pull(node) %>% unique,
  "Elephantina" = bsize.Atlantogenata.table %>% filter(label=="Elephantina") %>% pull(node) %>% unique,
  "Mammuthus" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
  "Proboscidae" = bsize.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(node) %>% unique,
  "Tethytheria" = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
  "Paenungulata" = bsize.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(node) %>% unique,
  "Pseudoungulata" = bsize.Atlantogenata.table %>% filter(label=="Pseudoungulata") %>% pull(node) %>% unique,
  "Afrotheria" = bsize.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
  "Afrosorcida" = bsize.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(node) %>% unique,
  "Macroscelidae" = bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
  "Xenarthra" = bsize.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(node) %>% unique,
  "Atlantogenata" = bsize.Atlantogenata.table %>% filter(label=="Atlantogenata") %>% pull(node) %>% unique
)

@@

bsize.Atlantogenata <- bsize.Atlantogenata.table %>% as.treedata()
```

```{r Body Size Data: All, include=F}
# Tree
st.innerNodes <- tribble(
  ~"Node",                                                     ~"label",
  MRCA(st.full, "Loxodonta.africana", "Loxodonta.cyclotis"), "Loxodontini",
  MRCA(st.full, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"), "Loxodona",
  MRCA(st.full, "Loxodonta.africana", "Elephas.maximus"), "Elephantidae",
  MRCA(st.full, "Elephas.maximus", "Mammuthus.primigenius"), "Elephantina",
  MRCA(st.full, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
  MRCA(st.full, "Loxodonta.africana", "Mammut.americanum"), "Proboscidae",
  MRCA(st.full, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
  MRCA(st.full, "Loxodonta.africana", "Procavia.capensis"), "Paenungulata",
  MRCA(st.full, "Loxodonta.africana", "Orycteropus.afer"), "Pseudoungulata",
  MRCA(st.full, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
  MRCA(st.full, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosorcida",
  MRCA(st.full, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
  MRCA(st.full, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), "Xenarthra",
  MRCA(st.full, "Dasypus.novemcinctus", "Loxodonta.africana"), "Atlantogenata"
)
st.full %<>% 
  as_tibble() %>% 
  select(-label.y, label=label.x)

@@

tree.Atlantogenata.nodes <- list(
  "Loxodonta.africana"= tree.Atlantogenata.table %>% filter(label=="Loxodonta africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= tree.Atlantogenata.table %>% filter(label=="Trichechus manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= tree.Atlantogenata.table %>% filter(label=="Procavia capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= tree.Atlantogenata.table %>% filter(label=="Orycteropus afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= tree.Atlantogenata.table %>% filter(label=="Choloepus hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = tree.Atlantogenata.table %>% filter(label=="Dasypus novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = tree.Atlantogenata.table %>% filter(label=="Chrysochloris asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = tree.Atlantogenata.table %>% filter(label=="Echinops telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = tree.Atlantogenata.table %>% filter(label=="Elephantulus edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = tree.Atlantogenata.table %>% filter(label=="Loxodonta cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = tree.Atlantogenata.table %>% filter(label=="Palaeoloxodon antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = tree.Atlantogenata.table %>% filter(label=="Elephas maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = tree.Atlantogenata.table %>% filter(label=="Mammuthus columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = tree.Atlantogenata.table %>% filter(label=="Mammuthus primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = tree.Atlantogenata.table %>% filter(label=="Mammut americanum") %>% pull(node) %>% unique,
  "Loxodontini" = tree.Atlantogenata.table %>% filter(label=="Loxodontini") %>% pull(node) %>% unique,
  "Loxodona" = tree.Atlantogenata.table %>% filter(label=="Loxodona") %>% pull(node) %>% unique,
  "Elephantidae" = tree.Atlantogenata.table %>% filter(label=="Elephantidae") %>% pull(node) %>% unique,
  "Elephantina" = tree.Atlantogenata.table %>% filter(label=="Elephantina") %>% pull(node) %>% unique,
  "Mammuthus" = tree.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
  "Proboscidae" = tree.Atlantogenata.table %>% filter(label=="Proboscidae") %>% pull(node) %>% unique,
  "Tethytheria" = tree.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
  "Paenungulata" = tree.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(node) %>% unique,
  "Pseudoungulata" = tree.Atlantogenata.table %>% filter(label=="Pseudoungulata") %>% pull(node) %>% unique,
  "Afrotheria" = tree.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
  "Afrosorcida" = tree.Atlantogenata.table %>% filter(label=="Afrosorcida") %>% pull(node) %>% unique,
  "Macroscelidae" = tree.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
  "Xenarthra" = tree.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(node) %>% unique,
  "Atlantogenata" = tree.Atlantogenata.table %>% filter(label=="Atlantogenata") %>% pull(node) %>% unique
)

@@

tree.Atlantogenata.table %<>% mutate(highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)}))

@@

In order to determine the cancer risk at each node, we used a simplified multistage cancer risk model for body size and lifespan [@ArmitageAndDol; @Peto2015; @Peto2015b]. We defined the

@@

<!--
## Retroduplication of EEF1A1 in the stem-lineage of Tehthytheria

>>>>

---
title: Pervasive duplication of tumor suppressor genes preceded parallel evolution of large bodied Atlantogenatans
author:
  - name: Juan Manuel Vazquez
    email: juanvazquez@uchicago.edu
    affiliation: University of Chicago
    corresponding: juanvazquez@uchicago.edu
  - name: Vincent J Lynch
    email: vjlynch@buffalo.edu
    affiliation: University at Buffalo, SUNY
address:
  - code: University of Chicago
    address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
  - code: University at Buffalo, SUNY
    address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
abstract: |
  Cancer is an intrinsic disease of multicellular organisms. Within a species, the size of an animal, - correlated with the individual's number of cells - and its lifespan - correlated with increasing cellular damage over time - positively correlate with the risk any individual has to form tumors. Between species, however, we do not observe any correlation between size, lifespan, and cancer, a phenomenon that referred to as _Peto's Paradox_. Elephants are a particularly interesting member of this class of paradoxical animals, since they are a set of large species deeply nested in a clade of smaller species, indicating a recent gain of size. Recent work has identified several individual cases of gene duplicates contributing to the increased cancer resistance of elephants, which suggests that duplication of tumor suppressor genes may play a more general role in mediating Peto's Paradox by increasing cancer resistance in large, long-lived species. 
  By using a Reciprocal Best-Hit BLAT search approach, we investigated copy numbers of all protein-coding genes in _Atlantogenatan_ genomes to see if there is any correlation between the copy number of duplicates and changes body size along the phylogenetic tree. From an initial set of 18,011 protein-coding genes in hg38, we identified a median of 13,880 genes in _Atlantogenatan_ genomes, of which a median of 940 genes are duplicated. We find that, just as body size fluctuates throughout _Atlantogenata_, genes involved in tumor suppressor pathways are also duplicated throughout the phylogenetic tree. Extant species of elephants, however, show active transcription of both canonical and duplicated copies of tummor suppressors that duplicated prior to and during their sudden increase in body size, suggesting that the duplication of tumor suppressor genes facilitates the evolution of increased body size by compensating for the increased cancer risk.

author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
bibliography: rbhb_paenungulates.bib
output: 
  workflowr::wflow_html: default
  rticles::plos_article: 
    includes:
      in_header: preamble.tex
  word_document: default
csl: plos.csl
editor_options:
  chunk_output_type: console
---

@@

    ```{r Libraries and functions, include = F}
knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, # fig.width=5, 
                      fig.retina = T, fig.align = "left")

@@

anage_full <- read_tsv("../data/AnAge/anage_build14.txt", col_types = list("Weaning (days)"=col_number(), "Weaning weight (g)"=col_number(), "References" = col_character())) %>% unite("label", Genus, Species, sep=".")
anage <- anage_full %>% 
  select(label,Family, "Maximum longevity (yrs)", "Adult weight (g)") %>% 
  mutate(lnSize.anage = sapply(
    `Adult weight (g)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  ),
  Lifespan = sapply(
    `Maximum longevity (yrs)`, 
    function(x){ifelse(is.na(x), NA, log(x))}
  )
  ) %>% 
  select(label, Family, lnSize.anage, Lifespan)
```
```{r Query List, include=F}
UP000005640.full <- UP000005640 %>% 
  filter(!is.na(Name))
UP000005640.genenames <- unique(UP000005640.full$Name)
query.binTable <- list(AvA = gene_list, UP000005640 = UP000005640.genenames) %>% 
  binTableFromList() %>% 
  left_join(., UP000005640, by=c("Gene"="Name")) %>% 
  mutate(
    "C.ORF" = str_detect(Gene, "^C[0-9XY]{1,2}orf\\d+"),  # No C-Orfs
    "LOC" = str_detect(Gene, "^LOC\\d+"),   # No uncharacterized loci
    "LINC" = str_detect(Gene, "^LINC\\d+"),   # No Long "Non-Coding" RNA
    "HLA" = str_detect(Gene, "^HLA|^HCP5|^CD74"), # No HLAs
    "Histone" = str_detect(Gene, "^H[1-4]|^HIST\\d"), # No Histones
    "ZincFinger" = str_detect(Gene, "^ZNF\\d+"),   # No Zinc Fingers
    "Riboprotein" = str_detect(Gene, "^RP[LS]\\d+"),   # No Riboproteins
    "OlfactoryReceptor" = str_detect(Gene, "^OR\\d+"),   # No Olfactory Receptors
    "Uncharacterized" = str_detect(header, "[uU]ncharacterized"),  # No Uncharacterized proteins
    "Putative" = str_detect(header, "[pP]utative"),  # No putative proteins
    "Viral" = str_detect(header, "[pP]olyprotein|ERVK|HERV"),  # No viral proteins
    "Fragment" = str_detect(header, "[Ff]ragment")  # No fragment proteins
  )
```
```{r Body Size Data: BSize only, include=F}
bsize.innerNodes <- tribble(
  ~"Node",                                                     ~"label",
  MRCA(bsize.tree, "Loxodonta.africana", "Loxodonta.cyclotis"), "Loxodontini",
  MRCA(bsize.tree, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"), "Loxodona",
  MRCA(bsize.tree, "Loxodonta.africana", "Elephas.maximus"), "Elephantidae",
  MRCA(bsize.tree, "Elephas.maximus", "Mammuthus.primigenius"), "Elephantina",
  MRCA(bsize.tree, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
  MRCA(bsize.tree, "Loxodonta.africana", "Mammut.americanum"), "Proboscidea",
  MRCA(bsize.tree, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
  MRCA(bsize.tree, "Loxodonta.africana", "Procavia.capensis"), "Paenungulata",
  MRCA(bsize.tree, "Loxodonta.africana", "Orycteropus.afer"), "Pseudoungulata",
  MRCA(bsize.tree, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
  MRCA(bsize.tree, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosoricida",
  MRCA(bsize.tree, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
  MRCA(bsize.tree, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), "Xenarthra",
  MRCA(bsize.tree, "Dasypus.novemcinctus", "Loxodonta.africana"), "Atlantogenata"
)
bsize.tree %<>% as_tibble()

@@

size.scale <- bsize.tree %>% as_tibble %>% pull(lnSize) %>% unique %>% as.numeric() %>% sort
size.scale.min <- size.scale[1]
size.scale.max <- size.scale[length(size.scale)]
get.size.color <- function(n, scale = size.scale){
  scale.colors = viridis::viridis(n=length(scale)) %>% 
    set_names(., as.character(scale))
  scale.colors[as.character(n)]
}
bsize.Atlantogenata.nodes <- list(
  "Loxodonta.africana"= bsize.Atlantogenata.table %>% filter(label=="Loxodonta africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= bsize.Atlantogenata.table %>% filter(label=="Trichechus manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= bsize.Atlantogenata.table %>% filter(label=="Procavia capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= bsize.Atlantogenata.table %>% filter(label=="Orycteropus afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= bsize.Atlantogenata.table %>% filter(label=="Choloepus hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = bsize.Atlantogenata.table %>% filter(label=="Dasypus novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = bsize.Atlantogenata.table %>% filter(label=="Chrysochloris asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = bsize.Atlantogenata.table %>% filter(label=="Echinops telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = bsize.Atlantogenata.table %>% filter(label=="Elephantulus edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = bsize.Atlantogenata.table %>% filter(label=="Loxodonta cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = bsize.Atlantogenata.table %>% filter(label=="Palaeoloxodon antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = bsize.Atlantogenata.table %>% filter(label=="Elephas maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = bsize.Atlantogenata.table %>% filter(label=="Mammut americanum") %>% pull(node) %>% unique,
  "Loxodontini" = bsize.Atlantogenata.table %>% filter(label=="Loxodontini") %>% pull(node) %>% unique,
  "Loxodona" = bsize.Atlantogenata.table %>% filter(label=="Loxodona") %>% pull(node) %>% unique,
  "Elephantidae" = bsize.Atlantogenata.table %>% filter(label=="Elephantidae") %>% pull(node) %>% unique,
  "Elephantina" = bsize.Atlantogenata.table %>% filter(label=="Elephantina") %>% pull(node) %>% unique,
  "Mammuthus" = bsize.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
  "Proboscidea" = bsize.Atlantogenata.table %>% filter(label=="Proboscidea") %>% pull(node) %>% unique,
  "Tethytheria" = bsize.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
  "Paenungulata" = bsize.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(node) %>% unique,
  "Pseudoungulata" = bsize.Atlantogenata.table %>% filter(label=="Pseudoungulata") %>% pull(node) %>% unique,
  "Afrotheria" = bsize.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
  "Afrosoricida" = bsize.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(node) %>% unique,
  "Macroscelidae" = bsize.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
  "Xenarthra" = bsize.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(node) %>% unique,
  "Atlantogenata" = bsize.Atlantogenata.table %>% filter(label=="Atlantogenata") %>% pull(node) %>% unique
)

@@

bsize.Atlantogenata <- bsize.Atlantogenata.table %>% as.treedata()
```
```{r Body Size Data: All, include=F}
# Tree
st.innerNodes <- tribble(
  ~"Node",                                                     ~"label",
  MRCA(st.full, "Loxodonta.africana", "Loxodonta.cyclotis"), "Loxodontini",
  MRCA(st.full, "Palaeoloxodon.antiquus", "Loxodonta.cyclotis"), "Loxodona",
  MRCA(st.full, "Loxodonta.africana", "Elephas.maximus"), "Elephantidae",
  MRCA(st.full, "Elephas.maximus", "Mammuthus.primigenius"), "Elephantina",
  MRCA(st.full, "Mammuthus.columbi", "Mammuthus.primigenius"), "Mammuthus",
  MRCA(st.full, "Loxodonta.africana", "Mammut.americanum"), "Proboscidea",
  MRCA(st.full, "Loxodonta.africana", "Trichechus.manatus"), "Tethytheria",
  MRCA(st.full, "Loxodonta.africana", "Procavia.capensis"), "Paenungulata",
  MRCA(st.full, "Loxodonta.africana", "Orycteropus.afer"), "Pseudoungulata",
  MRCA(st.full, "Loxodonta.africana", "Echinops.telfairi"), "Afrotheria",
  MRCA(st.full, "Chrysochloris.asiatica", "Echinops.telfairi"), "Afrosoricida",
  MRCA(st.full, "Elephantulus.edwardii", "Chambius.kasserinensis"), "Macroscelidae",
  MRCA(st.full, "Dasypus.novemcinctus", "Choloepus.hoffmanni"), "Xenarthra",
  MRCA(st.full, "Dasypus.novemcinctus", "Loxodonta.africana"), "Atlantogenata"
)
st.full %<>% 
  as_tibble() %>% 
  select(-label.y, label=label.x)

@@

tree.Atlantogenata.nodes <- list(
  "Loxodonta.africana"= tree.Atlantogenata.table %>% filter(label=="Loxodonta africana") %>% pull(node) %>% unique,
  "Trichechus.manatus"= tree.Atlantogenata.table %>% filter(label=="Trichechus manatus") %>% pull(node) %>% unique,
  "Procavia.capensis"= tree.Atlantogenata.table %>% filter(label=="Procavia capensis") %>% pull(node) %>% unique,
  "Orycteropus.afer"= tree.Atlantogenata.table %>% filter(label=="Orycteropus afer") %>% pull(node) %>% unique,
  "Dasypus.novemcinctus"= tree.Atlantogenata.table %>% filter(label=="Choloepus hoffmanni") %>% pull(node) %>% unique,
  "Choloepus.hoffmanni" = tree.Atlantogenata.table %>% filter(label=="Dasypus novemcinctus") %>% pull(node) %>% unique,
  "Chrysochloris.asiatica" = tree.Atlantogenata.table %>% filter(label=="Chrysochloris asiatica") %>% pull(node) %>% unique,
  "Echinops.telfairi" = tree.Atlantogenata.table %>% filter(label=="Echinops telfairi") %>% pull(node) %>% unique,
  "Elephantulus.edwardii" = tree.Atlantogenata.table %>% filter(label=="Elephantulus edwardii") %>% pull(node) %>% unique,
  "Loxodonta.cyclotis" = tree.Atlantogenata.table %>% filter(label=="Loxodonta cyclotis") %>% pull(node) %>% unique,
  "Palaeoloxodon.antiquus" = tree.Atlantogenata.table %>% filter(label=="Palaeoloxodon antiquus") %>% pull(node) %>% unique,
  "Elephas.maximus" = tree.Atlantogenata.table %>% filter(label=="Elephas maximus") %>% pull(node) %>% unique,
  "Mammuthus.columbi" = tree.Atlantogenata.table %>% filter(label=="Mammuthus columbi") %>% pull(node) %>% unique,
  "Mammuthus.primigenius" = tree.Atlantogenata.table %>% filter(label=="Mammuthus primigenius") %>% pull(node) %>% unique,
  "Mammut.americanum" = tree.Atlantogenata.table %>% filter(label=="Mammut americanum") %>% pull(node) %>% unique,
  "Loxodontini" = tree.Atlantogenata.table %>% filter(label=="Loxodontini") %>% pull(node) %>% unique,
  "Loxodona" = tree.Atlantogenata.table %>% filter(label=="Loxodona") %>% pull(node) %>% unique,
  "Elephantidae" = tree.Atlantogenata.table %>% filter(label=="Elephantidae") %>% pull(node) %>% unique,
  "Elephantina" = tree.Atlantogenata.table %>% filter(label=="Elephantina") %>% pull(node) %>% unique,
  "Mammuthus" = tree.Atlantogenata.table %>% filter(label=="Mammuthus") %>% pull(node) %>% unique,
  "Proboscidea" = tree.Atlantogenata.table %>% filter(label=="Proboscidea") %>% pull(node) %>% unique,
  "Tethytheria" = tree.Atlantogenata.table %>% filter(label=="Tethytheria") %>% pull(node) %>% unique,
  "Paenungulata" = tree.Atlantogenata.table %>% filter(label=="Paenungulata") %>% pull(node) %>% unique,
  "Pseudoungulata" = tree.Atlantogenata.table %>% filter(label=="Pseudoungulata") %>% pull(node) %>% unique,
  "Afrotheria" = tree.Atlantogenata.table %>% filter(label=="Afrotheria") %>% pull(node) %>% unique,
  "Afrosoricida" = tree.Atlantogenata.table %>% filter(label=="Afrosoricida") %>% pull(node) %>% unique,
  "Macroscelidae" = tree.Atlantogenata.table %>% filter(label=="Macroscelidae") %>% pull(node) %>% unique,
  "Xenarthra" = tree.Atlantogenata.table %>% filter(label=="Xenarthra") %>% pull(node) %>% unique,
  "Atlantogenata" = tree.Atlantogenata.table %>% filter(label=="Atlantogenata") %>% pull(node) %>% unique
)

@@

tree.Atlantogenata.table %<>% mutate(
  highlight = sapply(seq_along(label), function(x, l=species.highlight, Label=label, PhyloPicUID=phyloPicUID){ifelse(Label[x]%in% names(l), PhyloPicUID[x], NA)}),
  size.change = node %>% sapply(., branch.size.ratio, tree.data=tree.Atlantogenata.table)
  )

@@

In order to determine the cancer risk $K$ at each node, we first needed to calcultate ancestral lifespans at each node. To do so, we used a Phylogenetic Generalized Least-Square Regression (PGLS) [@Felsenstein1985; @MartinsAndHansen1997] to calculate estimated ancestral lifespans across _Atlantogenata_ using our estimates for body size at each node.

Next, we used a simplified multistage cancer risk model for body size $D$ and lifespan $t$: $K \approx Dt^6$ [@ArmitageAndDoll; @Peto2015; @Peto2015b]. To determine the change in $K$ between nodes, we obtained the ratio between the cancer risk $K_{1}$ at any given node, and the cancer risk $K_{2}$ at its ancestral node, using the equation $\frac{K_{2}}{K_{1}} \approx \frac{D_{1}t_{1}^6}{D_{2}t_{2}^6}$. Finally, to simplify comparisons, we calculated the fold change cancer risk between a node and its ancestor as $\log_{2}(\frac{K_{2}}{K_{1}})$.

@@
